---
title: "Data Exploration"
output:
  pdf_document: default
  html_document: default
date: "2024-09-12"
name: "Team 4"
---

```{r setup, include=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars, echo=FALSE, results='hide'}
summary(cars)
```

Loading the data - will be changed to the data from the drive

```{r, echo = FALSE, results='hide', message=FALSE, warning=FALSE}
setwd("C:/Users/lucil/Desktop/Master/MA/Dprep/Team 4 Project")
library(readr)
data_business <- read_csv("yelp_academic_dataset_business.csv")

```
# Summary Statistics 
### __Stars__ 

This section will explain the key statistics for the stars column as well as depict a plot of this column for a better understanding of the data. As our research will focus on the impact on these ratings, it is important to have a good understanding of this variable.  

```{r, echo = FALSE, results = 'hide'}
library(ggplot2)
summary(data_business$stars)
mean_stars <- mean(data_business$stars, na.rm = TRUE)
median_star <- median(data_business$stars, na.rm = TRUE)
min_starvalue <- min(data_business$stars, na.rm = TRUE)
max_starvalue <- max(data_business$stars, na.rm = TRUE)
rounded_mean_star <- round(mean_stars,2)

print(paste("Mean Star Rating:", mean_stars))
print(paste("Rounded Mean Star Rating:", rounded_mean_star))
print(paste("Median Star Rating:", median_star))
print(paste("Maximum Star Rating:", max_starvalue))
print(paste("Minimum Star Rating:",min_starvalue))
```
```{r, echo = FALSE}
# Create a data frame for structured output
summary_df <- data.frame(
  Statistic = c("Mean Star Rating", "Rounded Mean Star Rating", "Median Star Rating", "Maximum Star Rating", "Minimum Star Rating"),
  Value = c(mean_stars, rounded_mean_star, median_star, max_starvalue, min_starvalue)
)

# Print the summary table with better visuals
knitr::kable(summary_df, format = "markdown", caption = "Summary of Star Ratings")
```
```{r, echo=FALSE}
ggplot(data_business, aes(x = stars)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Star Ratings", x = "Star Ratings", y = "Frequency") +
  theme_minimal()
```

As depicted in the graph the most common rating obtained by the business on Yelp is of 4 stars. On the other hand, 1 star ratings are the least common. 

__Useful Information__

In the following section, when referring to the "Count" of a variable, it refers to the amount of times that this specific variable is present throughout the businesses in the Yelp data set. 

### __States__ 

This section will depict the location distribution of business among the different states in the USA as the Yelp Reviews are from these location. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
data_business$state <- as.factor(data_business$state)
ggplot(data_business, aes(x = state)) +
  geom_bar(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Count of Businesses by State", 
       x = "State", 
       y = "Count") +
  theme_minimal()
```

This figure allows us to better understand the geographical distribution of the businesses, which might of interest when assesing the reviews and ratings. 


### __Categories__ 


Only the top 20 categories are depicted in the following table for illustrative purposes. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(knitr)

# Split the categories and count the occurrences
category_counts <- data_business %>%
  mutate(categories = strsplit(as.character(categories), ", ")) %>% # Split by comma and space
  unnest(categories) %>% # Transform the list column into rows
  count(categories, sort = TRUE) 

# View the summary statistics
top_categories <- category_counts %>%
  top_n(20, n) %>%  
  arrange(desc(n))  

kable(top_categories, col.names = c("Category", "Count"),
      caption = "Count of Top 20 Categories", 
      format = "markdown")
```


The code to plot all business categories can be found below hiding. 

```{r, echo=FALSE, results='hide', include=FALSE}
library(ggplot2)

ggplot(category_counts, aes(x = reorder(categories, n), y = n)) +
  geom_bar(stat = "identity", fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Count of Categories", 
       x = "Categories", 
       y = "Count") +
  coord_flip() + 
  theme_minimal()
```

This figure represents the __Top 20 categories__ of businesses that appear more on Yelp. 

To obtain a better illustrative depiction of the categories only the 20 top categories are depicted on this plot.  

```{r, echo=FALSE}
top_categories <- category_counts %>%
  top_n(20, n) # You can also use slice_max(n, n = 20) from dplyr 1.0.0 or higher

ggplot(top_categories, aes(x = reorder(categories, n), y = n)) +
  geom_bar(stat = "identity", fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Top 20 Categories", 
       x = "Categories", 
       y = "Count") +
  coord_flip() + 
  theme_minimal()
```

### __Attributes__

This section explores in more detail the variable "Attributes" as it is one of the key elements for this research. As it is a key element in the research, the 30 most used attributes are depicted below. 

The variable attribute includes elements categorized as "True" or "False". If an element is indicated as "True" this means that said business has that attribute present while if it is indicated as "False" it indicates that that attribute is not present in that specific business. 

As we want to asses the impact of that the attributes have on ratings, firstly, the top 30 attributes present in most businesses will be depicted: 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(knitr)
library(stringr)

attribute_counts <- data_business %>%
  mutate(attributes = strsplit(as.character(attributes), ", ")) %>% 
  unnest(attributes) %>% 
  mutate(attributes = str_remove_all(attributes, "‘|’|\"|\\s*:\\s*False|\\s*:\\s*True")) %>% 
  filter(str_detect(attributes, "True")) %>% 
  count(attributes, sort = TRUE)

# View the summary statistics
top_attributes <- attribute_counts %>%
  top_n(30, n) %>%  
  arrange(desc(n))  

kable(top_attributes, col.names = c("Attributes", "Count"),
      caption = "Count of Top 30 Attributes", 
      format = "markdown")
``` 

As the lack of a an attribute can also have an impact on the rating of a business, the top 30 attributes less present on businesses will also be depicted below: 

```{r, echo=FALSE, warning=FALSE, message=FALSE} 

library(dplyr)
library(tidyr)
library(knitr)
library(stringr)

attribute_counts <- data_business %>%
  mutate(attributes = strsplit(as.character(attributes), ", ")) %>% 
  unnest(attributes) %>% 
  mutate(attributes = str_remove_all(attributes, "‘|’|\"|\\s*:\\s*False|\\s*:\\s*True")) %>% 
  filter(str_detect(attributes, "False")) %>% 
  count(attributes, sort = TRUE)

# View the summary statistics
top_attributes <- attribute_counts %>%
  top_n(30, n) %>%  
  arrange(desc(n))  

kable(top_attributes, col.names = c("Attributes", "Count"),
      caption = "Count of Top 30 Attributes", 
      format = "markdown")
``` 


```

