---
title: "Group Assignment Skills: Data Prep.&Workflow Mgt"
author: "Team 5"
date: "2024-09-06"
output: pdf_document
---

## Team 5
#Sophie van Hest
#Eveline Cai
#Mette Swanenberg
#Tyamo van der Ceelen

## Research Motivation

Our research question is: Is an individual’s fame related to his/her birth year?
By examining movie ratings and number of votes as a proxy for fame, this study seeks to explore whether people born in certain time periods are more likely to achieve fame in the film industry


## Data
Data1 includes:

- nconst (string) - alphanumeric unique identifier of the name/person
- primaryName (string)– name by which the person is most often credited
- birthYear – in YYYY format
- deathYear – in YYYY format if applicable, else '\\N'
- primaryProfession (array of strings)– the top-3 professions of the person
- knownForTitles (array of tconsts) – titles the person is known for

Data2 includes:

- tconst (string) - alphanumeric unique identifier of the title
- averageRating – weighted average of all the individual user ratings
- numVotes - number of votes the title has received

```{r}
url1 <- "https://datasets.imdbws.com/name.basics.tsv.gz"
url2 <- "https://datasets.imdbws.com/title.ratings.tsv.gz"

file1 <- "name.basics.tsv.gz"
file2 <- "title.ratings.tsv.gz"

if (!file.exists(file1)) {
  download.file(url1, file1)
}
if (!file.exists(file2)) {
  download.file(url2, file2)
}

data1 <- read.csv(gzfile(file1), sep = "\t")
data2 <- read.csv(gzfile(file2), sep = "\t")

head(data1)
head(data2)
```

```{r}
str(data1)
str(data2)
```
```{r}
summary(data1$birthYear)
```

```{r}
summary(data2$averageRating)
```
```{r}
summary(data2$numVotes)
```

## Data exploration

```{r}
#Libraries:
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
```
Plot of the distribution of birth years
```{r}
data1$birthYear <- as.numeric(data1$birthYear)
ggplot(data1, aes(x = birthYear)) +
  geom_histogram(binwidth = 10, fill = "lightblue", color = "black") +
  labs(title = "Distribution of Birth Years", x = "Birth Year", y = "Count") +
  theme_minimal() +
  scale_x_continuous(limits = c(1750, NA))
```
Plots of the distribution of average movie ratings and number of votes
```{r}
# Plotting distribution of average movie ratings
ggplot(data2, aes(x = averageRating)) +
  geom_histogram(binwidth = 0.5, fill = "lightgreen", color = "black") +
  labs(title = "Distribution of Average Movie Ratings", x = "Average Rating", y = "Count") +
  theme_minimal()

# Plotting distribution of number of votes
ggplot(data2, aes(x = numVotes)) +
  geom_histogram(binwidth = 0.1, fill = "lightpink", color = "black") +
  labs(title = "Distribution of Number of Votes", x = "Number of Votes", y = "Count") +
  theme_minimal() +
  scale_x_log10()
```














