---
title: "Data Exploration"
output:
  pdf_document: default
  html_document: default
date: "2024-09-12"
name: "Team 4"
params:
  suppress_messages: TRUE
---

```{r setup, include=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars, echo=FALSE, results='hide'}
```

```{r, echo = FALSE, results='hide', message=FALSE, warning=FALSE}
library(googledrive)

# Authenticate with Google Drive
drive_deauth()

# Folder ID from your Google Drive link
folder_id <- "1ioJVCsr5pJ5tAa2dPJ9yxIvL6rYmDSl1"pleas

# List all files in the folder
files_in_folder <- drive_ls(as_id(folder_id))

# Filter for the file named 'yelp_academic_dataset_business.csv'
csv_file <- files_in_folder[files_in_folder$name == "yelp_academic_dataset_business.csv", ]

# Check if the file exists before attempting download
if (nrow(csv_file) > 0) {
  # Download the CSV file to the working directory
  drive_download(as_id(csv_file$id), path = file.path(getwd(), "yelp_academic_dataset_business.csv"), overwrite = TRUE)
  cat("File downloaded successfully.")
} else {
  cat("The file 'yelp_academic_dataset_business.csv' was not found in the folder.")
}

data_business <- read_csv("yelp_academic_dataset_business.csv") 
```
# Summary Statistics 
### Variable: __Stars__ 

This section will explain the key statistics for the stars column as well as depict a plot of this column for a better understanding of the data. As our research will focus on the impact on these ratings, it is important to have a good understanding of this variable.  

```{r, echo = FALSE, results = 'hide'}
library(ggplot2)
summary(data_business$stars)
mean_stars <- mean(data_business$stars, na.rm = TRUE)
median_star <- median(data_business$stars, na.rm = TRUE)
min_starvalue <- min(data_business$stars, na.rm = TRUE)
max_starvalue <- max(data_business$stars, na.rm = TRUE)
rounded_mean_star <- round(mean_stars,2)

print(paste("Mean Star Rating:", mean_stars))
print(paste("Rounded Mean Star Rating:", rounded_mean_star))
print(paste("Median Star Rating:", median_star))
print(paste("Maximum Star Rating:", max_starvalue))
print(paste("Minimum Star Rating:",min_starvalue))
```
```{r, echo = FALSE}
# Create a data frame for structured output
summary_df <- data.frame(
  Statistic = c("Mean Star Rating", "Rounded Mean Star Rating", "Median Star Rating", "Maximum Star Rating", "Minimum Star Rating"),
  Value = c(mean_stars, rounded_mean_star, median_star, max_starvalue, min_starvalue)
)

# Print the summary table with better visuals
knitr::kable(summary_df, format = "markdown", caption = "Summary of Star Ratings")
```
```{r, echo=FALSE}
ggplot(data_business, aes(x = stars)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Star Ratings", x = "Star Ratings", y = "Frequency") +
  theme_minimal()
```

As depicted in the graph the most common rating obtained by the business on Yelp is of 4 stars. On the other hand, 1 star ratings are the least common. 

__Useful Information__

In the following section, when referring to the "Count" of a variable, it refers to the amount of times that this specific variable is present throughout the businesses in the Yelp data set. 

### Variable: __States__ 

This section will depict the location distribution of business among the different states in the USA as the Yelp Reviews are from these location. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
data_business$state <- as.factor(data_business$state)
ggplot(data_business, aes(x = state)) +
  geom_bar(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Count of Businesses by State", 
       x = "State", 
       y = "Count") +
  theme_minimal()
```

This figure allows us to better understand the geographical distribution of the businesses, which might of interest when assesing the reviews and ratings. 


### Variable: __Categories__ 


Only the top 20 categories are depicted in the following table for illustrative purposes. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(knitr)

# Split the categories and count the occurrences
category_counts <- data_business %>%
  mutate(categories = strsplit(as.character(categories), ", ")) %>% # Split by comma and space
  unnest(categories) %>% # Transform the list column into rows
  count(categories, sort = TRUE) 

# View the summary statistics
top_categories <- category_counts %>%
  top_n(20, n) %>%  
  arrange(desc(n))  

kable(top_categories, col.names = c("Category", "Count"),
      caption = "Count of Top 20 Categories", 
      format = "markdown")
```


The code to plot all business categories can be found below hiding. 

```{r, echo=FALSE, results='hide', include=FALSE}
library(ggplot2)

ggplot(category_counts, aes(x = reorder(categories, n), y = n)) +
  geom_bar(stat = "identity", fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Count of Categories", 
       x = "Categories", 
       y = "Count") +
  coord_flip() + 
  theme_minimal()
```

This figure represents the __Top 20 categories__ of businesses that appear more on Yelp. 

To obtain a better illustrative depiction of the categories only the 20 top categories are depicted on this plot.  

```{r, echo=FALSE}
top_categories <- category_counts %>%
  top_n(20, n) # You can also use slice_max(n, n = 20) from dplyr 1.0.0 or higher

ggplot(top_categories, aes(x = reorder(categories, n), y = n)) +
  geom_bar(stat = "identity", fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Top 20 Categories", 
       x = "Categories", 
       y = "Count") +
  coord_flip() + 
  theme_minimal()
```

### Variable: __Attributes__

This section explores in more detail the variable "Attributes" as it is one of the key elements for this research. As it is a key element in the research, the 30 most used attributes are depicted below. 

The variable attribute includes elements categorized as "True" or "False". If an element is indicated as "True" this means that said business has that attribute present while if it is indicated as "False" it indicates that that attribute is not present in that specific business. 

As we want to asses the impact of that the attributes have on ratings, firstly, the top 30 attributes present in most businesses will be depicted: 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(knitr)
library(stringr)

attribute_counts <- data_business %>%
  mutate(attributes = strsplit(as.character(attributes), ", ")) %>% 
  unnest(attributes) %>% 
  mutate(attributes = str_remove_all(attributes, "‘|’|\"|\\s*:\\s*False|\\s*:\\s*True")) %>% 
  filter(str_detect(attributes, "True")) %>% 
  count(attributes, sort = TRUE)

# View the summary statistics
top_attributes <- attribute_counts %>%
  top_n(30, n) %>%  
  arrange(desc(n))  

kable(top_attributes, col.names = c("Attributes", "Count"),
      caption = "Count of Top 30 Attributes", 
      format = "markdown")
``` 

As the lack of a an attribute can also have an impact on the rating of a business, the top 30 attributes less present on businesses will also be depicted below: 

```{r, echo=FALSE, warning=FALSE, message=FALSE} 

library(dplyr)
library(tidyr)
library(knitr)
library(stringr)

attribute_counts <- data_business %>%
  mutate(attributes = strsplit(as.character(attributes), ", ")) %>% 
  unnest(attributes) %>% 
  mutate(attributes = str_remove_all(attributes, "‘|’|\"|\\s*:\\s*False|\\s*:\\s*True")) %>% 
  filter(str_detect(attributes, "False")) %>% 
  count(attributes, sort = TRUE)

# View the summary statistics
top_attributes <- attribute_counts %>%
  top_n(30, n) %>%  
  arrange(desc(n))  

kable(top_attributes, col.names = c("Attributes", "Count"),
      caption = "Count of Top 30 Attributes", 
      format = "markdown")
``` 


```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(jsonlite)

# Function to parse the attribute string into a named list
parse_attributes <- function(attr_string) {
  # Clean the string to make it valid JSON syntax
  cleaned_string <- gsub("u'", "'", attr_string)  # Remove the 'u' prefix
  cleaned_string <- gsub("'", "\"", cleaned_string)  # Replace single quotes with double quotes
  cleaned_string <- gsub("\\\\\"", "\"", cleaned_string)  # Remove escape characters before quotes
  cleaned_string <- gsub('"(True|False)"', '\\L\\1', cleaned_string, perl = TRUE)  # Convert "True"/"False" to lowercase
  cleaned_string <- gsub(": None", ": null", cleaned_string)  # Replace None with null
  
  # Remove any invalid escaping inside JSON strings
  cleaned_string <- gsub('(?<=:)\\s*""(.*?)""', '"\\1"', cleaned_string, perl = TRUE)
  
  # Ensure the string is enclosed in curly braces
  cleaned_string <- paste0("{", str_remove_all(cleaned_string, "^\\{|\\}$"), "}")
  
  # Attempt to parse the string into a list
  parsed_list <- tryCatch(fromJSON(cleaned_string), error = function(e) NULL)
  
  return(parsed_list)
}

# Apply the parse_attributes function to each row and convert to a dataframe
parsed_attributes <- data_business %>%
  mutate(attributes_list = lapply(attributes, parse_attributes)) %>%
  unnest_wider(attributes_list)

# Display the resulting dataframe
parsed_attributes
```
```{r}

attributes <- data_business$attributes
```


### Variable: __Review Count__  
The purpose of this section is to explore the key statistics of the review count column as well as depicting plots of this column for a better understanding of the data. As our research will distinguish between businesses that have a low number of reviews versus businesses that have a high number of reviews, it is crucial to have a good understanding of this variable.


```{r, echo = FALSE, results = 'hide'}
library(ggplot2)
summary(data_business$review_count)
mean_review <- mean(data_business$review_count, na.rm = TRUE)
median_review <- median(data_business$review_count, na.rm = TRUE)
min_reviewvalue <- min(data_business$review_count, na.rm = TRUE)
max_reviewvalue <- max(data_business$review_count, na.rm = TRUE)
rounded_mean_review <- round(mean_review,2)

print(paste("Mean Review Count:", mean_review))
print(paste("Rounded Mean Review Count:", rounded_mean_review))
print(paste("Median Review Count:", median_review))
print(paste("Maximum Review Count:", max_reviewvalue))
print(paste("Minimum Review Count:",min_reviewvalue))
```
```{r, echo = FALSE}
# Create a data frame for structured output
summary_df <- data.frame(
  Statistic = c("Mean Review Count", "Rounded Mean Review Count", "Median Review Count", "Maximum Review Count", "Minimum Review Count"),
  Value = c(mean_review, rounded_mean_review, median_review, max_reviewvalue, min_reviewvalue)
)

# Print the summary table with better visuals
knitr::kable(summary_df, format = "markdown", caption = "Summary of Review Count")
```



The variable ‘review count’ can be divided into 4 quartiles with the following ranges:
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
quartiles <- quantile(data_business$'review_count', probs = c(0.25, 0.50, 0.75))

data_business <- na.omit(data_business)
quartile_boundaries <- quantile(data_business$review_count, probs = 0:4/4, na.rm = TRUE)
quartile_ranges <- data.frame(
  Quartile = c("1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile"),
  Lower_Bound = quartile_boundaries[1:4],
  Upper_Bound = quartile_boundaries[2:5]
)

print(quartile_ranges)

data_business <- na.omit(data_business)
data_business$quartile <- cut(data_business$review_count, 
                              breaks = quantile(data_business$review_count, probs = 0:4/4, na.rm = TRUE), 
                              include.lowest = TRUE, 
                              labels = c("1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile"))

```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(gt)

# Create a gt table from quartile_ranges
quartile_table <- quartile_ranges %>%
  gt() %>%
  tab_header(
    title = "Review Count Quartile Ranges"  # Title for the table
  ) %>%
  cols_label(
    Quartile = "Quartile",                   # Column Label for Quartile
    Lower_Bound = "Lower Bound",             # Column Label for Lower Bound
    Upper_Bound = "Upper Bound"              # Column Label for Upper Bound
  ) %>%
  fmt_number(
    columns = vars(Lower_Bound, Upper_Bound),  # Format number columns
    decimals = 0                               # Set decimal places
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "lightblue")  # Change background color for better visibility
    ),
    locations = cells_body(
      columns = everything()  # Apply to all columns
    )
  ) %>%
  tab_options(
    table.font.size = 12,                       # Font size
    table.font.color = "darkblue",               # Font color
    table.border.top.color = "black",           # Top border color
    table.border.bottom.color = "black"         # Bottom border color
  )

# Display the table
quartile_table
```

Each quartile represents a specific range of review counts, from the lowest 25% to the highest 25%. This segmentation helps us understand the distribution of review activity among businesses. The table depicted above implies that businesses that fall into the first quartile have between 5 and 8 reviews. Businesses that fall into the second quartile have between 8 and 15 reviews, businesses in the third quartile have between 15 and 37 reviews and businesses in the fourth quartile have between 37 and 7568 reviews. This is valuable information because it allows to distinguish between businesses that have a low review count versus businesses with a high review count.


```{r, echo=FALSE, warning=FALSE, message=FALSE}
install.packages("ggplot2")
library(ggplot2)

hist(data_business$review_count, 
     main = "Distribution of Ratings", 
     xlab = "Ratings", 
     col = "lightblue", 
     border = "black")
     
data_business$quartile <- cut(data_business$review_count, 
                              breaks = quantile(data_business$review_count, probs = 0:4/4, na.rm = TRUE), 
                              include.lowest = TRUE, 
                              labels = c("1st Quartile", "2nd Quartile", "3rd  
                            Quartile", "4th Quartile"))
                            
quartile_counts <- table(data_business$quartile)

barplot(quartile_counts, 
        main = "Businesses by Review Count Quartiles", 
        xlab = "Quartiles", 
        ylab = "Number of Businesses", 
        col = c("lightblue", "lightgreen", "lightcoral", "lightgoldenrod"), 
        border = "black")
```



The graph above depicts how many businesses fall into each “category” of review count that is elaborated above:

-	1st Quartile: 44228 businesses

-	2nd Quartile: 33646 businesses

-	3rd Quartile: 35303 businesses

-	4th Quartile: 37169 businesses



__Review Count and Stars__

In this analysis, we created quartiles based on the review count variable to categorize businesses into four distinct groups representing different levels of review activity. We then examined the distribution of the ‘stars’ variable across these quartiles. This involved plotting the number of businesses within each quartile against their corresponding star ratings.

The quartiles allowed us to segment businesses into groups ranging from the lowest to the highest 25% of review counts. By plotting star ratings against these quartiles, we were able to visually assess how star ratings vary across different levels of review activity. The resulting plots illustrate any patterns or trends in star ratings relative to review count quartiles. This approach provides insights into the relationship between the volume of reviews and the star ratings assigned to businesses. The graph and table below depict our findings:


```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE}

library(dplyr)
summary_data <- data_business %>%
  group_by(quartile, stars) %>%
  tally() %>%
  ungroup()

print(summary_data)
library(ggplot2)
ggplot(summary_data, aes(x = stars, y = n, fill = quartile)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of Star Ratings by Review Count Quartiles", 
       x = "Star Rating", 
       y = "Number of Businesses") +
  theme_minimal()
  
  print(summary_data, n = Inf)

print(summary_data, n = Inf)

```
```{r, echo=FALSE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

options(repos = c(CRAN = "https://cloud.r-project.org/"))
install.packages("knitr")
install.packages("gt")
library(gt)

gt_table <- summary_data %>%
  gt() %>%
  tab_header(
    title = "Summary of Businesses by Quartile and Star Rating")
library(knitr)

gt_table
```

```

