---
title: "Group Assignment Skills: Data Prep.&Workflow Mgt"
author: "Team 5"
date: "2024-09-06"
output:
  html_document:
    self_contained: TRUE
    output_dir: "../../gen/data-preparation/output/"
---

## Team 5
#Sophie van Hest
#Eveline Cai
#Mette Swanenberg
#Tyamo van der Ceelen

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```


## Initial Analyses
```{r}
library(dplyr)
library(readr)
library(ggplot2)
```

```{r include=FALSE}
data <- read_csv("../../data/data.csv")
```

Create control variable actorExperience: the number of movies an actor has been involved in
```{r}
data <- data %>%
  group_by(primaryName) %>%
  mutate(actorExperience = n()) %>%
  ungroup()
```

As R is running out of memory while trying to handle a large computation, likely due to the size and structure of our dataset we made a few changes.
Only include actors with at least 10 movies:
```{r}
data_filtered <- data %>%
  filter(actorExperience >= 10)
```
Only include movies with at least 100 number of votes:
```{r}
data_filtered <- data_filtered %>%
  filter(numVotes >= 50000)
```
Keep the top 5 genres, lump the other genres into 'other'. 
```{r}
#install.packages("forcats")
library(forcats)
data_filtered$genre <- fct_lump(data_filtered$genre, n = 5)
```
We reduced the magnitude of the startYear variable by grouping it into 10-year intervals (decades)
```{r}
data_filtered$startYear <- as.numeric(data_filtered$startYear)
data_filtered$startYear_decade <- (data_filtered$startYear %/% 10) * 10
```

Prepare dummy variables for genre, primaryName and startYear. We also drop all the rows that are missing the average Rating.
```{r}
data_filtered <- data_filtered %>%
  mutate(genre = factor(genre), primaryName = factor(primaryName), startYear_decade = factor(startYear_decade)) %>%
  tidyr::drop_na(averageRating)
```

Our regression model looks as follows:
Dependent variable: 
- averageRating (IMDb rating of the movie)

Independent variables:
- actorExperience (the number of films an actor has been in)
- numVotes (popularity of the movie)
- genre (movie genre)
- startYear_decade (the year the movie was made)


```{r}
summary(lm(averageRating ~ actorExperience + numVotes + genre + startYear_decade, data = data_filtered))
```
